### 2018-05-08 paper/paper.tex
abtract was hardcoded in paper.tex

\documentclass[sigconf, anonymous]{acmart}

\usepackage{xspace}
\usepackage{listings}
\usepackage{siunitx}

\newcommand{\todo}[1]{{\color{red}{\textbf{\em [TODO: #1]}}}\xspace}
\newcommand{\TODO}[1]{\todo{#1}}

\input{figures}

\fancyhf{} % Remove fancy page headers 
\fancyhead[C]{Paper \#XXX} % TODO: replace 9999 with your paper number
\fancyfoot[C]{\thepage}

\setcopyright{none} % No copyright notice required for submissions
%\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
\acmYear{2018}

\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions

%%\ccsPaper{9999} % TODO: replace with your paper number once obtained

\newcommand{\speculake}{ExSpectre\xspace}

\begin{document}
\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title

\input{abstract.tex}

% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10002978.10003029.10011703</concept_id>
%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
% -- end of section to replace with generated code

\keywords{Malware; Covert Channel; Weird Machines; TK} % TODO: replace with your keywords

\maketitle

\input{introduction.tex}

\input{background.tex}

\input{architecture.tex}

\input{payloads.tex}

\input{implementation.tex}

\input{discussion.tex}

\input{related-work.tex}

\input{conclusion.tex}

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\appendix
\input{appendix_spasm.tex}
\end{document}