### 2018-05-08 paper/conclusion.tex

\section{Conclusion}

In this paper we have presented \speculake, a model for hiding computation in
speculative execution that is fundamentally different than modern methods of
obfuscating computation. Through a series of experiments we have classified the
capabilities and limitations of this speculative primitive and demonstrated
various example applications. We have demonstrated the widely varying
capabilities of speculative execution, in particular showing that it allows for
arbitrary computation from a Turing Machine, emulation, transparent and remote
communication, and process level data encryption. As a result current analysis
techniques for detecting process cooperation, and describing the behavior of
program binaries are no longer definitively reliable.

Given the benefits that speculative execution provides to an attacker and the
ease of such an attack, the feasibility of an \speculake exploit is high and is
not limited to data side-channels or process isolation violations, as was
originally thought in the wake of Spectre. The speculative primitive developed
in this work not only provides a guarded form of computation, but also enables
cryptographic protection against current analysis and debugging techniques.

Software isolation techniques are critical in protecting against exploitation
models, however, complex optimization has often neglected these principles in
favor of performance. The stop-gap measures proposed in this work contain
security and performance trade-offs, though we advocate for a more robust
solution in future processor generations. Long term solutions may require that
we take a step backward before we can securely step forward.