### 2018-06-21 paper/conclusion.tex

\section{Conclusion}

%In this paper 
We have presented \speculake, a model for hiding computation in
speculative execution that is fundamentally different than existing methods of
code obfuscation. Through a series of experiments we have classified the
capabilities and limitations of this speculative primitive and demonstrated
various example applications. We have demonstrated the potential
of using speculative execution in several applications, including a Turing
machine, SPASM emulator, remotely-triggered payloads, and AES decryption.
Currently analysis techniques for reverse engineering are insufficient to reason
about the behavior of these programs.

Given the benefits that speculative execution provides to an attacker and the
ease of such an attack, the feasibility of an \speculake exploit is high and is
not limited to data side-channels or process isolation violations, as was
originally thought in the wake of Spectre. The speculative primitive developed
in this work not only provides a guarded form of computation, but also enables
cryptographic protection against current analysis and debugging techniques.

Ultimately, silicon and microarchitecture patches will be needed to secure CPUs
against this kind of malware. Meanwhile, attackers may
iterate and find new variants of \speculake-like malware.
In the meantime, new detection techniques and
software-level mitigations are desperately needed.

%Software isolation techniques are critical in protecting against exploitation
%models, however, complex optimization has often neglected these principles in
%favor of performance. The stop-gap measures proposed in this work contain
%security and performance trade-offs, though we advocate for a more robust
%solution in future processor generations. Long term solutions may require that
%we take a step backward before we can securely step forward.
%