### 2018-06-21 paper/introduction.tex

\section{Introduction}

Modern CPU designs use speculative execution to maintain high instruction
throughput, with the goal of improving performance. In speculative execution,
CPUs execute likely future instructions while they wait for other slower
instructions to complete. When the CPU's guess of future instructions is correct, the
benefit is faster execution performance. When its guess is wrong, the CPU
simply discards the speculated results and continues executing along the true path.

Previously, it was assumed that speculative execution results
remain invisible if discarded, as careful CPU design maintains strict separation
between speculative results and updates to architectual state.
However, recent research has revealed side channels that violate this
separation, and researchers have demonstrated ways to exfiltrate results from speculative
computation. Most notably, the Spectre vulnerability allows attackers to leak
information from purposefully mis-speculated branches in a victim
process~\cite{spectre}. The Meltdown vulnerability uses speculative results of an
unauthorized memory read to sidestep page faults and leak protected memory from
the kernel~\cite{meltdown}. Both of these vulnerabilities focus on extracting
secret data from a process or operating system.

\medskip

In this paper, we explore another attack enabled by speculative execution:
\speculake, which \emph{hides computation} within
the ``speculative world''. Taking advantage of the CPU's speculation to secretly
perform computation,
we can produce binaries that thwart existing reverse engineering
techniques. Because the speculative parts of a program never ``truly'' execute,
we can hide program functionality in the unreachable dead code in a program.
Even a full instruction trace, captured by a hardware debugger or software
emulator, will be unable to capture the logic performed speculatively.
This technique could lead to sophisticated malware that hides its behavior
from both static and dynamic analysis.

Existing malware use several techniques to evade detection and
make it difficult for analysts to determine payload behavior of reported malware. 
For example, binary \emph{packers} or \emph{crypters} encode an executable payload as
data that must be ``unpacked'' at runtime, making it difficult to tell
statically what a program will do~\cite{malware-packers}. Malware may also use
\emph{triggers} that only run the payload when certain conditions are present, preventing
it from executing when it is inside an analysis sandbox or 
debugger~\cite{balzarotti2010efficient,red-pill}.

However, with some effort, these existing malware techniques can be defeated.
Analysts can use
dynamic execution to unpack malware and reveal its
behavior~\cite{balzarotti2010efficient}, and can use symbolic execution or code
coverage fuzzers to determine the inputs or triggers that will reveal malicious
behavior~\cite{moser2007exploring,schwartz2010all,wang2017angr,egele2012survey}.

\speculake provides a new technique to malware authors, allowing them to hide
program functionality in code that does not execute at runtime.

At a high level, \speculake consists of two parts: a payload program, and a
trigger program. When run by itself, the payload program executes a pattern of
indirect jumps and measures a cache side channel. While the trigger program is
not running, the payload program is effectively inactive (and may optionally do
some decoy operation). When the trigger program runs, it executes a similar
pattern of indirect jumps (with similar source and destination addresses as the
payload program), effectively training the CPU's branch predictor to the jump
pattern specified by the trigger program.

Importantly, the trigger program and payload program's indirect jump patterns
diverge on the destination of their final jumps. However, because the trigger
program has trained the CPU's branch predictor, the CPU speculates that the
payload program will continue following the pattern of the trigger program,
causing it to execute a \emph{speculative gadget} in the payload
program. This execution path is shown in Figure~\ref{fig:high-level}.

We emphasize that this gadget is in a region that is neither
read nor executed by the payload program, and after the CPU discovers the
mis-speculation in the payload program, it will discard the results from the
speculative gadget and continue executing from the correct destination. However,
the payload program has a limited speculative window where it can perform
computation, and can communicate results back to the ``real world'' payload 
program via a cache side channel.

%Figure~\ref{fig:high-level}

\FigHighLevel

While it's possible for the trigger and payload programs to be bundled in a
single program, an analyst aware of \speculake could easily find the speculative
gadget in the payload program based on the jump pattern in the trigger program.
However, if the trigger and payload are kept separate, the analyst has a much
harder job and must identify both.
Moreover, it is possible that the trigger program be a
seemingly-unrelated benign program already on the victim's computer. We
investigate using the OpenSSL library as a benign trigger
program in Section~\ref{subsec:openssl}. If the trigger program is another benign
program on the system, the analyst has the difficult task of identifying which
program, library, or even operating system component is responsible for training
the CPU's branch predictor, and finding the specific set of jumps that occur at
runtime. To make matters worse, the payload program can include dummy jump
patterns that are effectively dead ends for the analyst, as they do not
ultimately call the hidden speculative gadget.

Simulating or modelling the speculative execution path is a difficult task for a
program analyst hoping to reverse engineer an \speculake binary. First, the
analyst must reverse engineer and accurately model the closed-source proprietary
components of the target CPU, including the branch predictor, cache hierarchy,
out-of-order execution, and hyperthreading. In contrast, the \speculake
author only has to use a partial model of these components and produce binaries
that take advantage of them, while the analyst's model must be complete to
capture all potential \speculake uses. Second, the analyst must run all
potential trigger programs through the simulator, including benign programs with
realistic inputs. Both of these contribute to a time-consuming and expensive
endeavor for would-be analysts.

% Results summary
In order to study the potential of \speculake, we implement several example
payload programs and evaluate their performance. We find that the speculative
window is limited by the reorder buffer and physical register file sizes, which
allows us to execute between 100 and 200 instructions speculatively on modern
Intel CPUs. While brief, we are able to perform execution in short steps,
communicating intermediate results back to the ``real world'' part of the
payload program. Using this technique, we demonstrate implementing a universal
Turing machine (demonstrating arbitrary computation), a custom instruction set
architecture that fits within the constraints of speculative execution, and show
the ability to perform AES decryption of a text block via AES encrypt/decrypt
instructions.

Using these building blocks, we demonstrate the practicality of hiding arbitrary
computation by implementing a reverse shell in our speculative instruction set.
We show that this simple payload is able to perform several system calls in a
reasonable time, ultimately launching a dial-back TCP shell in just over
2~milliseconds after the trigger program starts.

%Using these building blocks, we demonstrate the practicality of hiding arbitrary
%computation by decrypting an AES-encrypted ARM binary in the speculative world
%one instruction at a time, returning the decrypted next instruction to the real
%world part of the payload program, which updates a virtual machine. We show that
%we are able to decrypt and process approximately 25 instructions per second.

% contributions?
% - explore the limits of speculative execution
% - propose novel concept of hiding computation in speculative execution
% - implement example applications using this concept, including decrypting
%   data speculatively
% - demonstrate using benign program (openssl) as trigger
% - identify defenses and discuss ways to counter them

% Move to discussion:
% TODO: we don't talk at all about how an analyst might try to enumerate
% the potential entry points to the speculative world, and how that can be
% made difficult

%%%%%%%%%%%%%%