### 2018-07-26 paper/paper.tex
update paper to use NDSS class

\documentclass[conference]{IEEEtran}

\usepackage{xspace}
\usepackage{listings}
\usepackage{siunitx}
\usepackage{balance}
\usepackage{graphicx}

\newcommand{\todo}[1]{{\color{red}{\textbf{\em [TODO: #1]}}}\xspace}
\newcommand{\TODO}[1]{\todo{#1}}

\input{figures}

% \fancyhf{} % Remove fancy page headers 
% \fancyhead[C]{Paper \#629} % TODO: replace 9999 with your paper number
% \fancyfoot[C]{\thepage}

% \setcopyright{none} % No copyright notice required for submissions
% %\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
% \acmYear{2018}

% \settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions

%%\ccsPaper{9999} % TODO: replace with your paper number once obtained

\newcommand{\speculake}{ExSpectre\xspace}

\begin{document}
\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
\author{\IEEEauthorblockN{Anonymized for Review}
% \IEEEauthorblockA{Georgia Institute of Technology\\
% someemail@somedomain.com}
% \and
% \IEEEauthorblockN{Homer Simpson}
% \IEEEauthorblockA{Twentieth Century Fox\\
% homer@thesimpsons.com}
% \and
% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
% \IEEEauthorblockA{Starfleet Academy\\
% someemail@somedomain.com}
}

\IEEEoverridecommandlockouts
\makeatletter\def\@IEEEpubidpullup{9\baselineskip}\makeatother
\IEEEpubid{\parbox{\columnwidth}{Permission to freely reproduce all or part
    of this paper for noncommercial purposes is granted provided that
    copies bear this notice and the full citation on the first
    page. Reproduction for commercial purposes is strictly prohibited
    without the prior written consent of the Internet Society, the
    first-named author (for reproduction of an entire paper only), and
    the author's employer if the paper was prepared within the scope
    of employment.
}
\hspace{\columnsep}\makebox[\columnwidth]{}}

% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10002978.10003029.10011703</concept_id>
%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
% -- end of section to replace with generated code

% \keywords{Malware; Covert Channel; Weird Machines; Spectre} % TODO: replace with your keywords

\maketitle
\input{abstract}

\input{introduction.tex}

\input{background.tex}

\input{architecture.tex}

\input{payloads.tex}

\input{implementation.tex}

\input{discussion.tex}

\input{related-work.tex}

\input{conclusion.tex}

\bibliographystyle{IEEEtranS}
\bibliography{paper}

\appendix
\input{appendix_spasm.tex}
\end{document}