### 2018-08-02 paper/payloads.tex
first pass on turing and decrypt

\section{Application Payloads}

While the amount of computation done in a single speculative execution is small,
we demonstrate several applications that can take advantage of multiple
speculative runs to carry out computation.

As a first step, we observe that the speculative primitive can be used to trivially
implement a finite state machine: logic can be done in the speculative
world, while updates to the state are communicated to the real world where they
are stored. On the next run of the speculative instructions, the state is read from the
real world state (along with any inputs), state transitions are computed and communicated back.
In this mode, the state is maintained by the real world,
while updates are controlled by code executed speculatively.

We further observe we are not limited to finite state machines, but can support
any model of computation where \emph{updates} to any state are finite (i.e. can fit
within the bandwidth constraints of the speculative primitive). This
encompases Turing machines~\cite{turing1937computable} as well as certain random access machines, which we
investigate next.
Figure~\ref{fig:general_model} demonstrates the execution flow of a sample \speculake malware.

%Next, we observe that the instructions to be executed speculatively can be
%obscured further by encrypting them. This encryption allows for hidden
%computation, instructions are stored in the real world encrypted, they are then
%decrypted in the speculative world and passed to the real world to be executed.
%This method makes it appear as though the program is generating its own code.

\FigGeneralModel

\subsection{Turing Machine}
\label{subsec:turing}
To demonstrate that arbitrary computation can be performed cooperatively between
the speculative world and real world,
we implement a Turing machine, and configure it to run a 5-state Busy Beaver 
function~\cite{turing1937computable,chaitin1987computing,herken1992universal}. This configuration allows
us to run a large number of steps with very minimal logic.
%The Busy Beaver is a configuration which writes the most
%1s on the tape before halting with a limited number of states.

Updates to this Turing machine are computed speculatively, while the real
world keeps track of the state and full tape of the machine. Thus, the logic of
the machine is entirely contained in the speculative world, while the state may
be externally visible (e.g. to a dynamic debugger). We note that the machine
only operates when the trigger executes, making it difficult for an analyst with
only access to the Turing machine to determine exactly what the machine will do
from its initial state.

However, this toy example is meant only as an illustrative example of arbitrary
computation, notas a robust means of obfuscation. Indeed, even the initial state
of a Turing machine alone may reveal a significant amount of information.
Furthermore, the analyst may attempt to locate potential speculative entry
points, even without the help of the trigger program. We describe ways to address
both of these next.

\subsection{Unpacking and Decryption}
\label{subsec:decryption}

While a Turing machine demonstrates that arbitrary speculative computation
is possible, hiding malware this way has several drawbacks. First, Turing
machines are a poor choice for practical computation, as they are inefficient
and have no direct way to interface with the rest of the system (e.g. via system
calls). Second, as mentioned, they leave a great deal of information available
to the analyst, including the initial tape state, and a potentially small
(enumerable) number of possible speculative entry points.

We explore a more practical application of using \speculake to perform
\emph{decryption} speculatively. To hide keys from the analyst, the key and decryption
code only occur in the speculative world, while the initial payload program
contains only the ciphertext. While partial plaintext will be available in the
real world during execution, we emphasize that this only occurs when the trigger
runs. Before this, the state of the program reveals only the ciphertext that
will be decrypted. While the speculative entry point enumeration attack could be
used to reveal the keys used to decrypt this ciphertext, we describe a way to
derive the decryption key entirely from the trigger program. Thus, an analyst
that only has access to the payload program will be unable to learn the key or
decrypt the embedded ciphertext.

We also note that even when the trigger runs, decryption does not occur outside
the speculative context, meaning that any traditional traps or debugging
breakpoints placed on decryption instructions or routines will not occur, even
as they are used speculatively. These instructions could even be obfuscated themselves by
placing them in other misaligned instructions, and choosing a speculative entry
point that jumps to the middle of other instructions.

We note that 200 instructions is too short for most software-implemented
cryptography. However, modern Intel CPUs provide hardware support for AES, which
we find only takes a handful of $\mu$-ops to perform the instructions needed in
AES decryption. We discuss details of our speculative AES decryption in
Section~\ref{subsec:impl-aes}.

\medskip

% Could cut
%We leverage this support in Section~\ref{subsec:decryption} to
%increase the security of speculative decryption keys.
%
%There are numerous challenges for a reverse engineer to overcome to learn how a
%malicious program with this primitive works. As with previous examples, the
%malicious instructions are only executed when the correct trigger program is
%running (priming the jump table predictor \textbf{TK: word for this?}). However,
%when the to-be-executed instructions are executed the reverse engineer cannot
%locate these instructions in the ELF file, or generated source code. These
%instructions (when they even appear) seem to be generated by themselves, with no
%apparent cause. Additionally, there is no indicator that these instructions are
%encrypted as the decryption is done in the speculative world, meaning that no
%decryption (or key interaction) is done in a way visible to a reverse engineer.
%
%This method allows for arbitrary computation that is unobservable to any observer (static or dynamic).

% 
% Argument for obfuscating keyschedule
% 
% Challenge for Rev-Engineer
%   - Find the obfuscated key schedule
%   - Find the correct trigger program

\subsection{Virtual Machines}
We demonstrate in Section~\ref{subsec:turing} that general purpose computation 
is possible in the speculative environment in spite of the constraints 
imposed. However, a Turing Machine
is not an efficient or usable model for computation when
attempting to accomplish tasks on modern processors.

Using emulators in the ``real world'' allows for binary programs to be
placed encrypted and hidden in dead code and, as demonstrated in
Section~\ref{subsec:decryption}.
Traditional reverse engineering methods will reveal only that
emulation is being done, while the program being emulated remains encrypted.
Even when the trigger is running, only the parts of the code that execute would
be revealed to a careful analyst observing the CPU's committed state, and the
remainder of the emulated program would remain hidden.

We design a custom emulator and instruction set---SPASM (Speculative
Assembly)---that accommodates the constraints of the speculative primitive.
SPASM is a 6-bit Instruction set, where all instructions (including operand,
registers, and arguments) fit within 6-bits. This allows each step of the
speculative world to emit a single SPASM instruction to the real world for
emulation by a light-weight SPASM emulator. Using SPASM, developers can write
programs, assemble and encrypt them into a payload program. When the
associated trigger program runs, the payload will decrypt SPASM instructions in
the speculative world, and execute them one at a time.

% The assembly level abstraction that SPASM provides is still not extremely
% user friendly as it does not allow an author to write their encrypted 
% emulator payload in a high level language.

While the custom emulator that we developed gives higher level abstraction to
an author, it still requires programs to be written in a custom assembly language. 
We note that the \speculake model is not intrinsically linked to the SPASM emulator. 
A wrapper could be implemented around other existing emulators to construct
instructions incrementally through the fixed-width channel (e.g. using 4 8-bit
reads to reveal a single 32-bit ARM instruction).
This would allow for encrypted payloads to be written in higher level languages
like C or C++, and compiled to any target architecture the real-world emulator
can execute.