### 2018-08-08 paper/conclusion.tex
conclusion pass

\section{Conclusion}

%In this paper 
We have presented \speculake, a model for hiding computation in speculative
execution that is fundamentally different than existing methods of code
obfuscation. Through a series of experiments we have classified the capabilities
and limitations of this speculative primitive and demonstrated various example
applications. We have demonstrated the potential of using speculative execution
in several applications, including a Turing machine, SPASM emulator,
remotely-triggered payloads, and AES decryption. We have also examined Intel's
responses to Spectre and Meltdown and noted how their defenses affect \speculake
and how further variants of Spectre can be adapted to \speculake. Current
analysis techniques for reverse engineering are insufficient to reason about the
behavior of these programs.

%Given the benefits that speculative execution provides to an attacker and the
%ease of such an attack, the feasibility of an \speculake exploit is high and is
%not limited to data side-channels or process isolation violations, as was
%originally thought in the wake of Spectre. The speculative primitive developed
%in this work not only provides a guarded form of computation, but also enables
%cryptographic protection against current analysis and debugging techniques.

Ultimately, silicon and microarchitecture patches will be needed to secure CPUs
against this kind of malware. Meanwhile, attackers may
iterate and find new variants of \speculake-like malware.
In the meantime, new detection techniques and
software-level mitigations are desperately needed.

%Software isolation techniques are critical in protecting against exploitation
%models, however, complex optimization has often neglected these principles in
%favor of performance. The stop-gap measures proposed in this work contain
%security and performance trade-offs, though we advocate for a more robust
%solution in future processor generations. Long term solutions may require that
%we take a step backward before we can securely step forward.
%