### 2018-11-29 paper/paper.tex
fix formatting

\documentclass[conference]{IEEEtran}
\pagestyle{plain}

\usepackage{xspace}
\usepackage{listings}
\usepackage{siunitx}
\usepackage{balance}
\usepackage{graphicx}

\newcommand{\todo}[1]{{\color{red}{\textbf{\em [TODO: #1]}}}\xspace}
\newcommand{\TODO}[1]{\todo{#1}}

\input{figures}

% \fancyhf{} % Remove fancy page headers 
% \fancyhead[C]{Paper \#629} % TODO: replace 9999 with your paper number
% \fancyfoot[C]{\thepage}

% \setcopyright{none} % No copyright notice required for submissions
% %\acmConference[Anonymous Submission to ACM CCS 2018]{ACM Conference on Computer and Communications Security}{Due 08 May 2018}{Toronto, Canada}
% \acmYear{2018}

% \settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions

%%\ccsPaper{9999} % TODO: replace with your paper number once obtained

\newcommand{\speculake}{ExSpectre\xspace}

\begin{document}
\title{\speculake: Hiding Malware in Speculative Execution} % TODO: replace with your title
\author{\IEEEauthorblockN{Jack Wampler}
\IEEEauthorblockA{University of Colorado Boulder\\
jack.wampler@colorado.edu}
\and
\IEEEauthorblockN{Ian Martiny}
\IEEEauthorblockA{University of Colorado Boulder\\
ian.martiny@colorado.edu}
\and
\IEEEauthorblockN{Eric Wustrow}
\IEEEauthorblockA{University of Colorado Boulder\\
ewust@colorado.edu}
% \IEEEauthorblockA{Georgia Institute of Technology\\
% someemail@somedomain.com}
% \and
% \IEEEauthorblockN{Homer Simpson}
% \IEEEauthorblockA{Twentieth Century Fox\\
% homer@thesimpsons.com}
% \and
% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
% \IEEEauthorblockA{Starfleet Academy\\
% someemail@somedomain.com}
}
\IEEEoverridecommandlockouts
\makeatletter\def\@IEEEpubidpullup{6.5\baselineskip}\makeatother
\IEEEpubid{\parbox{\columnwidth}{
    Network and Distributed Systems Security (NDSS) Symposium 2019\\
    24-27 February 2019, San Diego, CA, USA\\
    ISBN 1-891562-55-X\\
    https://dx.doi.org/10.14722/ndss.2019.23409\\
    www.ndss-symposium.org
}
\hspace{\columnsep}\makebox[\columnwidth]{}}

% TODO: replace this section with code generated by the tool at https://dl.acm.org/ccs.cfm

%\begin{CCSXML}
%<ccs2012>
%<concept>
%<concept_id>10002978.10003029.10011703</concept_id>
%<concept_desc>Security and privacy~Usability in security and privacy</concept_desc>
%<concept_significance>500</concept_significance>
%</concept>
%</ccs2012>
%\end{CCSXML}

%\ccsdesc{Security and privacy~Use https://dl.acm.org/ccs.cfm to generate actual concepts section for your paper}
% -- end of section to replace with generated code

% \keywords{Malware; Covert Channel; Weird Machines; Spectre} % TODO: replace with your keywords

\maketitle
\input{abstract}

\input{introduction.tex}

\input{background.tex}

\input{architecture.tex}

\input{payloads.tex}

\input{triggers}

\input{implementation.tex}

\input{discussion.tex}

\input{related-work.tex}

\input{conclusion.tex}

\bibliographystyle{IEEEtranS}
\bibliography{paper}

\appendix
\input{appendix_spasm.tex}
\end{document}